Bayesian Sparsification of Deep $\cplx$-valued networks

### 2020-03-23 paper/paper.tex
thorough proofread 7 (sec 5.3 appendix D) * modified positioning in minst figures * moved CIFAR10 from appendix into main text, dropped old 2x replicated results * properly cite CIFAR10

% a couple of plots from MNIST and Fashion-MNIST
\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__cmp__VD__mnist__fft__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__VD__mnist__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy curve (VD, \texttt{fft}, MNIST):
    $2 \real / \cplx$ (\textit{top}) and $\real / \cplx$ (\textit{bottom}).
  }
  \label{fig:mnist-like__trade-off__fft}
\end{figure}

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__ARD__mnist__raw__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__cmp__ARD__mnist__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy curve (ARD, \texttt{raw}, MNIST):
    $\real / \cplx$ (\textit{top}) and $\real / \tfrac12 \cplx$ (\textit{bottom}).
  }
  \label{fig:mnist-like__trade-off__raw}
\end{figure}

@@

We use accuracy as the performance metric for multi-class image classification on CIFAR10 datasets.

% mention the same experimental setup as in MNIST
Experiments on CIFAR10 dataset are set up similarly to MNIST-like experiment, except
that we study VGG16 model \citep{simonyan_very_2015} and its $\cplx$-counterpart and
replicate each experiment only twice.
%
% describe the VGG16 model and its cplx-valued modification
The $\cplx$-VGG16 model is architecturally identical to VGG16, except for all layers being
replaced by their $\cplx$-valued versions. We halve the sizes of intermediate layers in
$\cplx$ model to assess comparability \citep{monning_evaluation_2018}.
%
Each network is trained stagewise for $20$, $40$,
and $20$ epochs on the full training split of $50k$ colour $32\times32$ images. We use
only raw colour features (\texttt{raw}). We apply random crop and horizontal flips as
data augmentation and use a learning rate scheduler in SGD, which scales the base learning
rate after $5$, $10$, and $20$-th epoch by $\tfrac1{10}$, $\tfrac1{20}$ and $\tfrac1{100}$,
respectively.

% draw some conclusions
Figure~\ref{fig:figure__cifar10__trade-off} shows that although the prior used in ARD method
offers slightly less compression, it makes up for it by equally slightly better accuracy.
Furthermore, the results imply that that the halved $\cplx$ network visibly underperforms
in comparison with its $\real$ counterpart, although in MNIST-like experiment $\cplx$-valued
network with the same inner feature dimensions demonstrates trade-off very similar to
$\real$-network, which has half the number of real parameters than the former.

% compare compressed cplx-vgg against real-vgg
\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__cifar__trade-off/legacy__augmentedcifar10__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    Compression-accuracy frontier for $\real$ and $\tfrac12 \cplx$ versions of VGG16 on CIFAR10.
  }
  \label{fig:figure__cifar10__trade-off}
\end{figure}

@@

\clearpage

\section{CIFAR10 experiments} % (fold)
\label{sec:cifar_experiments}

In addition to the results reported in the main paper, we conduct additional experiment
on CIFAR10 and VGG16 with a much more slowly annealed learning rate. Similar to MNIST-like
experiments, we start with the rate set to $10^{-3}$ and then reduce it by $10$ after the
$10$-the epoch. In summary, for the CIFAR10 experiments we reuse the set-up with the following
changes:
\begin{itemize}
  \item we allocate $20$, $40$, and $20$ epochs to each stage
  \item the weight of the KL divergence term varies over a smaller grid $
    C \in \{
      \tfrac32 2^{-\tfrac{k}2} \colon k=7, \cdots, 15
    \}
  $
  \item we use full $50$k training split of the CIFAR10 dataset for training and raw color
  features only
  \item dataset is randomly augmented: every image within a batch of size $128$ is randomly
  cropped and flipped horizontally
\end{itemize}
Random cropping is done by zero-padding each side of a $32\times 32$ image by four pixels
and then a extracting a random $32\times 32$ patch from the $40\times 40$ intermediate image.

We experiment with $\real$ VGG16 network \citep{simonyan_very_2015} and its $\cplx$-valued
variant, which straightforwardly replaces $\real$ layers with their $\cplx$ counterparts.

\begin{figure}[htb]
  \centering
  \includegraphics[width=\columnwidth]{figure__cifar__trade-off/appendix__augmentedcifar10__raw__-0.5.pdf}
  \caption{%
    The compression-accuracy trade-off for $\real$ and $\cplx$ VGG16 using raw input features.
  }
  \label{fig:appendix__cifar__trade-off__VGG16__raw}
\end{figure}

% section cifar_experiments (end)

>>>>

% a couple of plots from MNIST and Fashion-MNIST
\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__cmp__VD__mnist__fft__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__VD__mnist__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy curve (VD, \texttt{fft}, MNIST):
    $2 \real / \cplx$ (\textit{top}) and $\real / \cplx$ (\textit{bottom}).
  }
  \label{fig:mnist-like__trade-off__fft}
\end{figure}

\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__ARD__mnist__raw__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__cmp__ARD__mnist__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy curve (ARD, \texttt{raw}, MNIST):
    $\real / \cplx$ (\textit{top}) and $\real / \tfrac12 \cplx$ (\textit{bottom}).
  }
  \label{fig:mnist-like__trade-off__raw}
\end{figure}

@@

We conduct an experiment on the CIFAR10 classification dataset \citep{krizhevsky_learning_2009}
comprising $32\times 32$ color images of $10$ classes to study the compression-performance
trade-off of a moderately large deep network.
% same experimental setup as in MNIST, describe the VGG16 model and its cplx-valued modification
The experiment is set up similarly to MNIST-like experiment, except that we train $\real$
VGG16 network \citep{simonyan_very_2015} and its $\cplx$ variant, in which we have replaced
$\real$-valued layers with their $\cplx$-valued counterparts. $\real$ and $\cplx$ VGG16 have
the same feature sizes in intermediate layers.

% differences with mnist
Unlike experiment in sec.~\ref{sub:mnist_like_datasets}, we consider the raw features only,
use full training split, and allocate $20$, $40$, and $20$ epochs to each stage. During training
the mini-batch of $128$ samples is randomly augmented by randomly flipping horizontally and
cropping. Random cropping is done by zero-padding each side by four pixels and extracting a
random $32\times 32$ patch from the $40\times 40$ intermediate image.
%
We vary $C$ in \eqref{eq:elbo_with_coef} over $
  \{\tfrac32 2^{-\tfrac{k}2} \colon k=7, \cdots, 15\}
$, compare VD and ARD methods, and measure accuracy on CIFAR10 test split.

% compare compressed cplx-vgg against real-vgg
\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    % \includegraphics[width=\columnwidth]{figure__cifar__trade-off/legacy__augmentedcifar10__raw__-0.5.pdf}
    \includegraphics[width=\columnwidth]{figure__cifar__trade-off/appendix__augmentedcifar10__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy frontier for VGG16 ($\real$ and $\cplx$) on CIFAR10.
  }
  \label{fig:figure__cifar10__trade-off}
\end{figure}

% draw some conclusions
Figure~\ref{fig:figure__cifar10__trade-off} shows that, although ARD method offers slightly
less compression, it makes up for it by marginally better accuracy post fine-tune. VGG16 and
$\cplx$-VGG16 exhibit similar declining compression-accuracy trade-off, but due to higher
capacity $\cplx$-valued network yields higher accuracy \citep{monning_evaluation_2018}. Unlike
MNIST, fine-tuning stage significantly improves the test performance, which, we speculate, might
be due to dataset augmentation and deeper model.

@@
